### 2019-03-10 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/Ketipov_Kostadinov_Petrov_Zankinski_Balabanov.tex
Initial revision.

\documentclass[11pt]{article}
\usepackage{url}
\usepackage{graphicx,DCCN2019_en}

\pagestyle{fancy}
\fancyhead{}
\fancyfoot{}

\usepackage[utf8]{inputenc}
\linespread{1.0}

\usepackage{amsmath}
\usepackage{placeins}

\makeatletter
\fancyhead[RO]{\small DCCN 2019 \\ {23-27 September 2019}}
\fancyhead[LO]{\small Ketipov, Kostadinov, Petrov, Zankinski, Balabanov \\ Human-Computer MDC for TS Forecasting}
% Rumen Ketipov
% Georgi Kostadinov
% Plamen Petrov
% Iliyan Zankinski
% Todor Balabanov

\c@page=1

\makeatother

\title{Human-Computer Mobile Distributed Computing for Time Series Forecasting}

\author[1]{\small R.R. Ktipov}
\author[1]{\small G.B. Kostadinov}
\author[1]{\small P.D. Petrov}
\author[1]{\small \\I.A. Zankinski}
\author[0000-0003-3139-069X]{\small T.D. Balabanov}

\affil[1]{\footnotesize Institute of Information and Communication Technologies \break Bulgarian Academy of Sciences \break acad. Georgi Bonchev Str., block 2, office 514, 1113 Sofia, Bulgaria}

\email{rketipov@iit.bas.bg, g.kostadinov@iit.bas.bg, p.petrov@iit.bas.bg, iliyan@hsi.iccs.bas.bg, todorb@iinf.bas.bg}

\begin{document}

\udc{004.93}

{\let\newpage\relax\maketitle}

\vskip -1.5em

\footnotetext{This work was supported by private funding of Velbazhd Software LLC.}

\begin{abstract}

\keywords{distributed computing, time series forecasting, artificial neural networks, evolutionary algorithms}
\end{abstract}

\section{Introduction} \label{Introduction}

\section{Model Proposition} \label{Model Proposition}

\section{Experiments and Results} \label{Experiments and Results}

\section{Conclusion} \label{Conclusion}

\begin{thebibliography}{99}
\end{thebibliography}

\end{document}

### 2019-03-10 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/abstract.tex
Initial revision.

\documentclass[12pt,a4paper]{article}

\title{\bf Human-Computer Mobile Distributed Computing for Time Series Forecasting}

\author{Rumen Ketipov, Georgi Kostadinov, Plamen Petrov, \\
Iliyan Zankinski, Todor Balabanov\textsuperscript{[0000-0003-3139-069X]} \\
\\
Institute of Information and Communication Technologies \\
Bulgarian Academy of Sciences \\
\\
acad. Georgi Bonchev Str, block 2, office 514 \\
1113 Sofia, Bulgaria \\
\\
todorb@iinf.bas.bg \\
http://www.iict.bas.bg/}

\date{}

\begin{document}

\maketitle

Financial time series forecasting is an attractive and intensively researched area. One of the most interesting research directions in this field is the usage of artificial neural networks combined with evolutionary algorithms. In some studies evolutionary algorithms are used for artificial neural network topology optimization, but in other studies evolutionary algorithms are used for artificial neural network weights optimization. When evolutionary algorithms are used for weights optimization it is very common the training to be combined with back-propagation algorithm. Even when such hybrid training algorithm is used it is in the class of supervised training. Training examples are fed to the artificial neural network and its weights are modified according the selected training algorithm. Evolutionary algorithms have a common advantage compared to the exact numerical methods and it is the possibility for parallel calculations. When the parallel calculations are done on separate mobile devices it is the case of mobile distributed computing. It is very common the financial forecasting to be done by the usage of the past values in the time series, but it can be combined with human evaluation of the future values. In this research human-computer based mobile distributed computing solution is proposed. The users are using Android application, which trains artificial neural network with evolutionary algorithms and back-propagation, but users are capable to vote for increase or decrease of the future values. The most successful forecast votes of the users are used in artificial neural network training process.

\end{document}

### 2019-03-31 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/Ketipov_Kostadinov_Petrov_Zankinski_Balabanov.tex
Original template used in 2019 was provided.
Paper abstract was done.

@@
 \documentclass[11pt]{article}
+\usepackage[english]{babel}
 \usepackage{url}
@@
 \usepackage{amsmath}
-\usepackage{placeins}
 
@@
 
-\footnotetext{This work was supported by private funding of Velbazhd Software LLC.}
+\footnotetext{This work was supported with private funding by Velbazhd Software LLC.}
 
 \begin{abstract}
+Distributed computing became very popular in the last two decades.
+In many cases distributed computing projects are based on a donated calculation power.
+The most famous donated distributed computing project is SETI@home, which is related to deep space signals processing in attempt to find an alien life.
+The usage of donated distributed computing had its influence in the time series forecasting in the face of MoneyBee\cite{bohn01} project.
+MoneyBee project was a desktop screensaver application, which was calculating financial time series forecasting by training of artificial neural networks with evolutionary algorithms.
+With the expansion of the mobile devices in last decade it become relevant some donated distributed computing solutions to be developed as mobile applications.
+Such solution was developed at IICT-BAS\cite{tomov01}, which is based on Android Live Wallpaper technology.
+This research proposes an extension of the work done at IICT-BAS in the direction of human-computer based distributed computing by providing software capabilities of the users to vote for future financial changes.
 
@@
 
+\vskip 1.5em
+
 \begin{thebibliography}{99}
+
+\bibitem{bohn01} Bohn, A., Guting, T., Mansmann, T., \textbf{\textit{MoneyBee: A new product to predict stock market developments using artificial intelligence and increased calculation capacitiy}} (in German), et al.
+Wirtschaftsinf, vol.
+45, no.
+3, p.
+325--333, 2003.
+
+\bibitem{tomov01} Tomov, P., Zankinski, I., Barova, M., \textbf{\textit{Mobile Alternative of the MoneyBee Project for Financial Forecasting}}, Proceedings of the Annual University Scientific Conference of the National Military University Vasil Levski, Veliko Tarnovo, p.
+1085--1089, 2018.
+
 \end{thebibliography}

### 2019-03-31 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/Ketipov_Kostadinov_Petrov_Zankinski_Balabanov.tex
Small part of the introduction was added.

@@
 
+Financial time series forecasting is an attractive and intensively researched area.
+One of the most interesting research directions in this field is related to the usage of artificial neural networks combined with evolutionary algorithms.
+In some studies evolutionary algorithms are used for artificial neural network topology optimization, but in other studies evolutionary algorithms are used for artificial neural network weights optimization.
+When evolutionary algorithms are used for weights optimization it is very common the training to be combined with back-propagation algorithm.
+Even when such hybrid training algorithm is used it is in the class of supervised training.
+Training examples are fed to the artificial neural network and its weights are modified according the selected training algorithm.
+Evolutionary algorithms have a common advantage compared to the exact numerical methods and it is the possibility for parallel calculations.
+When the parallel calculations are done on separate mobile devices it is the case of mobile distributed computing.
+It is very common the financial forecasting to be done by the usage of the past values in the time series, but it can be combined with human evaluation of the future values.
+In this research human-computer based mobile distributed computing solution is proposed.
+The users are using Android application, which trains artificial neural network with evolutionary algorithms and back-propagation, but users are capable to vote for increase or decrease of the future values.
+The most successful forecast votes of the users are used in artificial neural network training process.
+
 \section{Model Proposition} \label{Model Proposition}

### 2019-04-17 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/Ketipov_Kostadinov_Petrov_Zankinski_Balabanov.tex
Small change in the phisical address.
Most of the paper was written.

@@
 \usepackage{graphicx,DCCN2019_en}
+\usepackage{placeins}
 
@@
 
-\author[1]{\small R.R. Ktipov}
+\author[1]{\small R.R. Ketipov}
 \author[1]{\small G.B. Kostadinov}
@@
 
-\affil[1]{\footnotesize Institute of Information and Communication Technologies \break Bulgarian Academy of Sciences \break acad. Georgi Bonchev Str., block 2, office 514, 1113 Sofia, Bulgaria}
+\affil[1]{\footnotesize Institute of Information and Communication Technologies \break Bulgarian Academy of Sciences \break acad. Georgi Bonchev Str., block 2, 1113 Sofia, Bulgaria}
 
@@
 
-Financial time series forecasting is an attractive and intensively researched area.
-One of the most interesting research directions in this field is related to the usage of artificial neural networks combined with evolutionary algorithms.
-In some studies evolutionary algorithms are used for artificial neural network topology optimization, but in other studies evolutionary algorithms are used for artificial neural network weights optimization.
+Financial time series forecasting is an attractive and intensively researched area \cite{nava01}.
+One of the most interesting research directions in this field is related to the usage of artificial neural networks combined with evolutionary algorithms \cite{zhang01}.
+In some studies evolutionary algorithms are used for artificial neural network topology optimization, but in other studies evolutionary algorithms are used for artificial neural network weights optimization \cite{aljarah01}.
 When evolutionary algorithms are used for weights optimization it is very common the training to be combined with back-propagation algorithm.
@@
 Training examples are fed to the artificial neural network and its weights are modified according the selected training algorithm.
-Evolutionary algorithms have a common advantage compared to the exact numerical methods and it is the possibility for parallel calculations.
+Evolutionary algorithms have a common advantage compared to the exact numerical methods and it is the possibility for parallel calculations \cite{altinoz01}.
 When the parallel calculations are done on separate mobile devices it is the case of mobile distributed computing.
@@
 
+MoneyBee screensaver (Fig.
+\ref{fig01}) was organized as desktop application, which runs in background when the computer is in idle mode.
+Artificial neural networks were trained with the usage of evolutionary algorithms.
+The main advantage of this approach is that it is using donated computation power of many participants.
+In the same time the biggest disadvantage is that the screensaver is used relatively rare only when the user is not using the computer.
+The other disadvantage is that personal computers are not used 24/7.
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{fig01.png}
+\centering
+\caption{MoneyBee screensaver.} \label{fig01}
+\end{figure}
+\FloatBarrier
+
+These disadvantages are addressed in the VitoshaTrade project (Fig.
+\ref{fig02}), developed in IICT-BAS.
+This implementation is done under Android operating system as active wallpaper.
+Inside the forecasting module multilayer perceptron is used.
+The training algorithm is combination between back-propagation and differential evolution.
+All calculations are done as active Android application, which works 24/7 in a background mode.
+When better artificial neural network weights are found communication with a remote server is done.
+Active wallpapers are used to visualize information on a full screen behind all other visual components.
+
+\begin{figure}
+\includegraphics[width=1.0\textwidth]{fig02.png}
+\centering
+\caption{VitoshaTrade active wallpaper.} \label{fig02}
+\end{figure}
+\FloatBarrier
+
+In both projects only pure mathematical information is used for prediction calculations.
+This limitation was overcome in ElectricSheep\cite{draves01} project.
+In this project rating system is used in which users vote for different fractal generated animations.
+Voting is done in two directions thumb up and thumb down.
+Each vote increase or decrease individual rating, which is directly related to its survival chances.
+The rating is not constant and it goes down with the time passing.
+
+This study proposes combination of the projects described above.
+VitoshaTrade project is extended in the direction of voting system addition.
+The Android operating system offers software modules called widgets.
+Widgets have graphical user interface and they occupy small part of the visible working area.
+Most commonly widgets are used for representation of small pieces of information, such astronomical time, weather forecast, calendar and many others.
+Widgets are much more interactive than the active wallpaper.
+The user is capable to input small pieces of information as voting for example.
+The other better advantage of the widgets is that users have much better control over visualization and placement of the widgets.
+In this study composite widget is proposed, which shows the thicker of the currency pair (EUR/USD in the example provided), value of the rate between the two currencies and arrows for vote up or vote down.
+
+Voting of the users is collected on PHP/MySQL based web server\cite{tomov01}, part of the server side implementation of the VitoshaTrade project.
+On the server side additional analysis is done.
+Votes which were more accurate are better taken for next generations in the global distributed differential evolution population.
+
 \section{Experiments and Results} \label{Experiments and Results}
 
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{fig03.png}
+\centering
+\caption{Forex EUR/USD time series.} \label{fig03}
+\end{figure}
+\FloatBarrier
+
 \section{Conclusion} \label{Conclusion}
+
+Pure technical analysis in time series field has its limits.
+Values measured in time are just one projection of the process, which goes under it.
+Any computer based calculation nowadays is not capable to act as human intuition.
+The mechanisms behind human intuition are not investigated enough and are still very difficult to recreate in the modern computers.
+Because of these reasons the proposed financial time series forecasting solution advances the pure technical analysis with minimal influence of human impact.
+The proposed simple voting system put in one place the power of machine learning algorithms and the human sixth sense feeling in an efficient and cost effective financial forecasting infrastructure.
+
+As future research it will be interesting different communication capabilities to be investigated as described in \cite{alexandrov01}.
+Also it will be interesting generalized artificial neural networks \cite{tashev01} to be used instead of multilayer perceptron.
 
@@
 
+\bibitem{nava01} Nava, N., Di Matteo, T., Aste, T., \textbf{\textit{Financial Time Series Forecasting Using Empirical Mode Decomposition and Support Vector Regression}}, RISKS, \textbf{6}(1), article 7, 2018.
+
+\bibitem{zhang01} Zhang, R., Tao, J., \textbf{\textit{A Nonlinear Fuzzy Neural Network Modeling Approach Using an Improved Genetic Algorithm}}, IEEE Transactions on Industrial Electronics, \textbf{65}(7), p.
+5882--5892, 2018.
+
+\bibitem{aljarah01} Aljarah, I., Faris, H., Mirjalili, S., \textbf{\textit{Optimizing connection weights in neural networks using the whale optimization algorithm}}, Soft Computing, Springer Berlin Heidelberg, \textbf{22}(1), p.
+1–15, 2016.
+
+\bibitem{altinoz01} Altinoz, O.T., Deb, K., \textbf{\textit{Late parallelization and feedback approaches for distributed computation of evolutionary multi-objective optimization algorithms}}, Neural Computing and Applications, Springer London, \textbf{30}(3), p.
+723--733, 2016.
+
 \bibitem{bohn01} Bohn, A., Guting, T., Mansmann, T., \textbf{\textit{MoneyBee: A new product to predict stock market developments using artificial intelligence and increased calculation capacitiy}} (in German), et al.
-Wirtschaftsinf, vol.
-45, no.
-3, p.
+Wirtschaftsinf, \textbf{45}(3), p.
 325--333, 2003.
@@
 
+\bibitem{draves01} Draves, S., \textbf{\textit{The Electric Sheep Screen-Saver: A Case Study in Aesthetic Evolution}}, Proceedings of Applications of Evolutionary Computing: EvoWorkkshops: EvoBIO, EvoCOMNET, EvoHOT, EvoIASP, EvoMUSART, and EvoSTOC Lausanne, LNCS 3449, p.
+458--467, 2005.
+
+\bibitem{alexandrov01} Alexandrov, A., \textbf{\textit{Comparative analysis of IEEE 802.15.4 based communication protocols used in wireless intelligent sensor systems}}, Proceedings of the International conference RAM, p.
+51--54, 2014.
+
+\bibitem{tashev01} Tashev, T., Hristov, H., \textbf{\textit{Modeling of synthesis of information processes with generalized nets}}, Cybernetics and Information Technologies, \textbf{3}(2), p.
+92--104, 2003.
+
 \end{thebibliography}

### 2019-04-18 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/Ketipov_Kostadinov_Petrov_Zankinski_Balabanov.tex
Initial full text of the paper.

@@
 One of the most interesting research directions in this field is related to the usage of artificial neural networks combined with evolutionary algorithms \cite{zhang01}.
-In some studies evolutionary algorithms are used for artificial neural network topology optimization, but in other studies evolutionary algorithms are used for artificial neural network weights optimization \cite{aljarah01}.
+In some studies evolutionary algorithms are used for artificial neural network topology optimization \cite{kapanova01}, but in other studies evolutionary algorithms are used for artificial neural network weights optimization \cite{aljarah01}.
 When evolutionary algorithms are used for weights optimization it is very common the training to be combined with back-propagation algorithm.
@@
 
-\section{Experiments and Results} \label{Experiments and Results}
+\section{Experiments} \label{Experiments}
+
+Experiments are done with FOREX financial time series of EUR/USD currency pair (Fig.
+\ref{fig03}).
+Time series is disassembled in training examples (7 values for lag and 3 values for lead).
+Back-propagation combined with differential evolution training of multilayer perceptron is done.
+Training process is continuous, which means that forecasting is done parallel with the training.
+Voting of the user directly influence the fitness value of the individuals in the distributed differential evolution population.
+Some users are better in their vision and are more successful during the voting process.
+Because of this tracking of the users is of great importance.
+More reliable votes are taken with higher coefficient for the further predictions.
+According some regulations, like GDPR in EU \cite{hristov01}, special care should be taken for the personal information collected on the servers side.
 
@@
 
+\bibitem{kapanova01} Kapanova, K., Dimov, I., Sellier, J.M., \textbf{\textit{A genetic approach to automatic neural network architecture optimization}}, Neural Computing and Applications, Springer London, \textbf{29}(5), p.
+1481--1492, 2016.
+
 \bibitem{aljarah01} Aljarah, I., Faris, H., Mirjalili, S., \textbf{\textit{Optimizing connection weights in neural networks using the whale optimization algorithm}}, Soft Computing, Springer Berlin Heidelberg, \textbf{22}(1), p.
@@
 
+\bibitem{hristov01} Hristov, P., Dimitrov, W., \textbf{\textit{The blockchain as a backbone of GDPR compliant frameworks}}, Proceedings of 8th International Multidisciplinary Symposium - Challenges and Opportunities For Sustainable Development Through Quality and Innovation in Engineering and Research Management, \textbf{20}(1), p.
+305--310, 2018.
+
 \bibitem{draves01} Draves, S., \textbf{\textit{The Electric Sheep Screen-Saver: A Case Study in Aesthetic Evolution}}, Proceedings of Applications of Evolutionary Computing: EvoWorkkshops: EvoBIO, EvoCOMNET, EvoHOT, EvoIASP, EvoMUSART, and EvoSTOC Lausanne, LNCS 3449, p.

### 2019-04-25 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/Ketipov_Kostadinov_Petrov_Zankinski_Balabanov.tex
Corrections done by Rumen.

@@
 
-\affil[1]{\footnotesize Institute of Information and Communication Technologies \break Bulgarian Academy of Sciences \break acad. Georgi Bonchev Str., block 2, 1113 Sofia, Bulgaria}
+\affil[1]{\footnotesize Institute of Information and Communication Technologies (IICT) \break Bulgarian Academy of Sciences (BAS) \break acad. Georgi Bonchev Str., block 2, 1113 Sofia, Bulgaria}
 
@@
 In many cases distributed computing projects are based on a donated calculation power.
-The most famous donated distributed computing project is SETI@home, which is related to deep space signals processing in attempt to find an alien life.
+One of the most famous donated distributed computing project is SETI@home, which is related to deep space signals processing in attempt to find an alien life.
 The usage of donated distributed computing had its influence in the time series forecasting in the face of MoneyBee\cite{bohn01} project.
 MoneyBee project was a desktop screensaver application, which was calculating financial time series forecasting by training of artificial neural networks with evolutionary algorithms.
-With the expansion of the mobile devices in last decade it become relevant some donated distributed computing solutions to be developed as mobile applications.
-Such solution was developed at IICT-BAS\cite{tomov01}, which is based on Android Live Wallpaper technology.
+With the expansion of the mobile devices in the last decade it become relevant some donated distributed computing solutions to be developed as mobile applications.
+Such a solution was developed at IICT-BAS\cite{tomov01}, which is based on Android Live Wallpaper technology.
 This research proposes an extension of the work done at IICT-BAS in the direction of human-computer based distributed computing by providing software capabilities of the users to vote for future financial changes.
@@
 When evolutionary algorithms are used for weights optimization it is very common the training to be combined with back-propagation algorithm.
-Even when such hybrid training algorithm is used it is in the class of supervised training.
+Even when such a hybrid training algorithm is used it is in the class of supervised training.
 Training examples are fed to the artificial neural network and its weights are modified according the selected training algorithm.
-Evolutionary algorithms have a common advantage compared to the exact numerical methods and it is the possibility for parallel calculations \cite{altinoz01}.
+Evolutionary algorithms have a common advantage compared to the exact numerical methods and that is the possibility for parallel calculations \cite{altinoz01}.
 When the parallel calculations are done on separate mobile devices it is the case of mobile distributed computing.
-It is very common the financial forecasting to be done by the usage of the past values in the time series, but it can be combined with human evaluation of the future values.
-In this research human-computer based mobile distributed computing solution is proposed.
-The users are using Android application, which trains artificial neural network with evolutionary algorithms and back-propagation, but users are capable to vote for increase or decrease of the future values.
+It is very common the financial forecasting to be done by the usage of the past values in the time series, but it can be also combined with human evaluation of the future values.
+In this research a human-computer based mobile distributed computing solution is proposed.
+The users are using an Android application which trains artificial neural network with evolutionary algorithms and back-propagation, but users are capable to vote for increase or decrease of the future values.
 The most successful forecast votes of the users are used in artificial neural network training process.
@@
 MoneyBee screensaver (Fig.
-\ref{fig01}) was organized as desktop application, which runs in background when the computer is in idle mode.
+\ref{fig01}) was organized as a desktop application, which runs in background when the computer is in idle mode.
 Artificial neural networks were trained with the usage of evolutionary algorithms.
 The main advantage of this approach is that it is using donated computation power of many participants.
-In the same time the biggest disadvantage is that the screensaver is used relatively rare only when the user is not using the computer.
+In the same time the biggest disadvantage is that the screensaver is used relatively rare, only when the user is not using the computer.
 The other disadvantage is that personal computers are not used 24/7.
@@
 \centering
-\caption{MoneyBee screensaver.} \label{fig01}
+\caption{MoneyBee screensaver} \label{fig01}
 \end{figure}
@@
 \ref{fig02}), developed in IICT-BAS.
-This implementation is done under Android operating system as active wallpaper.
-Inside the forecasting module multilayer perceptron is used.
-The training algorithm is combination between back-propagation and differential evolution.
+The implementation is done under an Android operating system as active wallpaper.
+Inside the forecasting module a multilayer perceptron is used.
+The training algorithm is a combination between back-propagation and differential evolution.
 All calculations are done as active Android application, which works 24/7 in a background mode.
-When better artificial neural network weights are found communication with a remote server is done.
+When better artificial neural network weights are found a communication with a remote server is established.
 Active wallpapers are used to visualize information on a full screen behind all other visual components.
@@
 \centering
-\caption{VitoshaTrade active wallpaper.} \label{fig02}
+\caption{VitoshaTrade active wallpaper} \label{fig02}
 \end{figure}
@@
 
-In both projects only pure mathematical information is used for prediction calculations.
+In both projects only a pure mathematical information is used for prediction calculations.
 This limitation was overcome in ElectricSheep\cite{draves01} project.
-In this project rating system is used in which users vote for different fractal generated animations.
-Voting is done in two directions thumb up and thumb down.
-Each vote increase or decrease individual rating, which is directly related to its survival chances.
+In this project it is used a rating system whereat the users vote for different fractal generated animations.
+Voting is done in two directions - thumb up and thumb down.
+Each vote increases or decreases individual rating, which is directly related to its survival chances.
 The rating is not constant and it goes down with the time passing.
 
-This study proposes combination of the projects described above.
-VitoshaTrade project is extended in the direction of voting system addition.
+This study proposes a combination of the projects described above.
+The VitoshaTrade project is extended in the direction of voting system addition.
 The Android operating system offers software modules called widgets.
@@
 The other better advantage of the widgets is that users have much better control over visualization and placement of the widgets.
-In this study composite widget is proposed, which shows the thicker of the currency pair (EUR/USD in the example provided), value of the rate between the two currencies and arrows for vote up or vote down.
+In this study a composite widget is proposed, which shows the thicker of the currency pair (EUR/USD in the example provided), the value of the rate between the two currencies, and arrows for vote up or vote down.
 
-Voting of the users is collected on PHP/MySQL based web server\cite{tomov01}, part of the server side implementation of the VitoshaTrade project.
+Voting of the users is collected on PHP/MySQL based web server\cite{tomov01}, a part of the server side implementation of the VitoshaTrade project.
 On the server side additional analysis is done.
@@
 Time series is disassembled in training examples (7 values for lag and 3 values for lead).
-Back-propagation combined with differential evolution training of multilayer perceptron is done.
-Training process is continuous, which means that forecasting is done parallel with the training.
-Voting of the user directly influence the fitness value of the individuals in the distributed differential evolution population.
-Some users are better in their vision and are more successful during the voting process.
-Because of this tracking of the users is of great importance.
+A back-propagation combined with differential evolution training of multilayer perceptron is done.
+The training process is continuous, which means that the forecasting is done parallel with the training.
+The voting of the user directly influences the fitness value of the individuals in the distributed differential evolution population.
+Some users are better in their vision and are more successful during the voting process, therefore the tracking of the users is of great importance.
 More reliable votes are taken with higher coefficient for the further predictions.
-According some regulations, like GDPR in EU \cite{hristov01}, special care should be taken for the personal information collected on the servers side.
+According to some regulations, like GDPR in EU \cite{hristov01}, a special care should be taken for the personal information collected on the servers side.
 
@@
 Pure technical analysis in time series field has its limits.
-Values measured in time are just one projection of the process, which goes under it.
+Values measured in time are just one projection of the process which goes under it.
 Any computer based calculation nowadays is not capable to act as human intuition.
@@
 Because of these reasons the proposed financial time series forecasting solution advances the pure technical analysis with minimal influence of human impact.
-The proposed simple voting system put in one place the power of machine learning algorithms and the human sixth sense feeling in an efficient and cost effective financial forecasting infrastructure.
+The proposed simple voting system consolidated the power of machine learning algorithms and the human sixth sense feeling in an efficient and cost effective financial forecasting infrastructure.
 
-As future research it will be interesting different communication capabilities to be investigated as described in \cite{alexandrov01}.
+As a future research it will be interesting different communication capabilities to be investigated as described in \cite{alexandrov01}.
 Also it will be interesting generalized artificial neural networks \cite{tashev01} to be used instead of multilayer perceptron.

### 2019-09-05 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/paper.tex
Initial Springer revision.

--- 2019-04-25 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/Ketipov_Kostadinov_Petrov_Zankinski_Balabanov.tex
@@
-\documentclass[11pt]{article}
-\usepackage[english]{babel}
-\usepackage{url}
-\usepackage{graphicx,DCCN2019_en}
+\documentclass[runningheads]{llncs}
+
+\usepackage{graphicx}
 \usepackage{placeins}
-
-\pagestyle{fancy}
-\fancyhead{} 
-\fancyfoot{}
-
-\usepackage[utf8]{inputenc}
-\linespread{1.0}
-
-\usepackage{amsmath}
-
-\makeatletter
-\fancyhead[RO]{\small DCCN 2019 \\ {23-27 September 2019}}
-\fancyhead[LO]{\small Ketipov, Kostadinov, Petrov, Zankinski, Balabanov \\ Human-Computer MDC for TS Forecasting}
-% Rumen Ketipov
-% Georgi Kostadinov
-% Plamen Petrov
-% Iliyan Zankinski
-% Todor Balabanov
-
-\c@page=1
-
-\makeatother
-
-\title{Human-Computer Mobile Distributed Computing for Time Series Forecasting}
-
-\author[1]{\small R.R. Ketipov}
-\author[1]{\small G.B. Kostadinov}
-\author[1]{\small P.D. Petrov}
-\author[1]{\small \\I.A. Zankinski}
-\author[0000-0003-3139-069X]{\small T.D. Balabanov}
-
-\affil[1]{\footnotesize Institute of Information and Communication Technologies (IICT) \break Bulgarian Academy of Sciences (BAS) \break acad. Georgi Bonchev Str., block 2, 1113 Sofia, Bulgaria}
-
-\email{rketipov@iit.bas.bg, g.kostadinov@iit.bas.bg, p.petrov@iit.bas.bg, iliyan@hsi.iccs.bas.bg, todorb@iinf.bas.bg}
 
@@
 
-\udc{004.93}
+\title{Human-Computer Mobile Distributed Computing for Time Series Forecasting\thanks{This work was supported with private funding by Velbazhd Software LLC.}}
 
-{\let\newpage\relax\maketitle}
+\author{
+	Rumen Ketipov \and
+	Georgi Kostadinov \and
+	Plamen Petrov \and \\
+	Iliyan Zankinskir \and
+	Todor Balabanov\orcidID{0000-0003-3139-069X}
+}
+% Rumen Ketipov - rketipov@iit.bas.bg
+% Georgi Kostadinov - g.kostadinov@iit.bas.bg
+% Plamen Petrov - p.petrov@iit.bas.bg
+% Iliyan Zankinski - iliyan@hsi.iccs.bas.bg
+% Todor Balabanov - todorb@iinf.bas.bg
 
-\vskip -1.5em
+\authorrunning{R. Ketipov et al.}
 
-\footnotetext{This work was supported with private funding by Velbazhd Software LLC.}
+\institute{
+	Institute of Information and Communication Technologies \\ 
+	Bulgarian Academy of Sciences \\ 
+	acad. Georgi Bonchev Str., block 2, 1113 Sofia, Bulgaria \\
+	\email{iict@bas.bg} \\  
+	\url{http://iict.bas.bg/}
+}
+
+\maketitle
 
@@
 
-\keywords{distributed computing, time series forecasting, artificial neural networks, evolutionary algorithms}
+\keywords{Distributed computing \and Time series forecasting \and Artificial neural networks \and Evolutionary algorithms.}
 \end{abstract}
@@
 
-\vskip 1.5em
-
-\begin{thebibliography}{99}
-
+\begin{thebibliography}{8}
 \bibitem{nava01} Nava, N., Di Matteo, T., Aste, T., \textbf{\textit{Financial Time Series Forecasting Using Empirical Mode Decomposition and Support Vector Regression}}, RISKS, \textbf{6}(1), article 7, 2018.
@@
 92--104, 2003.
-
 \end{thebibliography}

### 2019-09-08 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/paper.tex
Some fixes were done.

@@
 
-Financial time series forecasting is an attractive and intensively researched area \cite{nava01}.
-One of the most interesting research directions in this field is related to the usage of artificial neural networks combined with evolutionary algorithms \cite{zhang01}.
-In some studies evolutionary algorithms are used for artificial neural network topology optimization \cite{kapanova01}, but in other studies evolutionary algorithms are used for artificial neural network weights optimization \cite{aljarah01}.
+Financial time series forecasting is an attractive and intensively researched area\cite{nava01}.
+One of the most interesting research directions in this field is related to the usage of artificial neural networks combined with evolutionary algorithms\cite{zhang01}.
+In some studies evolutionary algorithms are used for artificial neural network topology optimization\cite{kapanova01}, but in other studies evolutionary algorithms are used for artificial neural network weights optimization\cite{aljarah01}.
 When evolutionary algorithms are used for weights optimization it is very common the training to be combined with back-propagation algorithm.
@@
 Training examples are fed to the artificial neural network and its weights are modified according the selected training algorithm.
-Evolutionary algorithms have a common advantage compared to the exact numerical methods and that is the possibility for parallel calculations \cite{altinoz01}.
+Evolutionary algorithms have a common advantage compared to the exact numerical methods and that is the possibility for parallel calculations\cite{altinoz01}.
 When the parallel calculations are done on separate mobile devices it is the case of mobile distributed computing.
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{fig01.png}
+\includegraphics[width=0.8\textwidth]{fig01.png}
 \centering
@@
 The rating is not constant and it goes down with the time passing.
+In the proposed model direction voting is implemented, just like it was done in ElectricSheep project.
+Users vote only up if they think that the price will go up or down if they think that the price will go down.
+There is no limit how many the user will press up or down button, just like it is implemented in ElectricSheep project.
+The intensity of the single user voting in one time interval (daily intervals for example) does not count.
+Only the first push of up/down button is taken in account.
+Users' voting is irregular, because in real-time operation the operator of the system has no control over donated human-computer participation.
 
 This study proposes a combination of the projects described above.
-The VitoshaTrade project is extended in the direction of voting system addition.
+The VitoshaTrade\cite{balabanov01} project is extended in the direction of voting system addition.
 The Android operating system offers software modules called widgets.
@@
 
-Experiments are done with FOREX financial time series of EUR/USD currency pair (Fig.
+Experiments are done with ForEx financial time series of EUR/USD currency pair (Fig.
 \ref{fig03}).
@@
 A back-propagation combined with differential evolution training of multilayer perceptron is done.
+As topology for the multilayer perceptron 7-5-3 is used implemented with Encog Machine Learning Framework\cite{heaton01} for Java.
+The input has 7 valuse, the output has 3 values and the hidden layer was seledted to be $5 = (7 + 3) / 2$.
 The training process is continuous, which means that the forecasting is done parallel with the training.
@@
 More reliable votes are taken with higher coefficient for the further predictions.
-According to some regulations, like GDPR in EU \cite{hristov01}, a special care should be taken for the personal information collected on the servers side.
+According to some regulations, like GDPR in EU\cite{hristov01}, a special care should be taken for the personal information collected on the servers side.
 
 \begin{figure}
-\includegraphics[width=0.8\textwidth]{fig03.png}
+\includegraphics[width=1.0\textwidth]{fig03.png}
 \centering
-\caption{Forex EUR/USD time series.} \label{fig03}
+\caption{ForEx EUR/USD time series.} \label{fig03}
 \end{figure}
 \FloatBarrier
+
+There is no limit how many people can participate in the proposed human-computer distributed computing system.
+The system itself is highly scalable, but experiments in this study were done in a closed group of the team which is working on the research.
 
@@
 Any computer based calculation nowadays is not capable to act as human intuition.
+Involving human intuition in the forecasting process is the main advantage of the proposed model.
 The mechanisms behind human intuition are not investigated enough and are still very difficult to recreate in the modern computers.
@@
 As a future research it will be interesting different communication capabilities to be investigated as described in \cite{alexandrov01}.
-Also it will be interesting generalized artificial neural networks \cite{tashev01} to be used instead of multilayer perceptron.
+Also it will be interesting generalized artificial neural networks\cite{tashev01} to be used instead of multilayer perceptron.
 
@@
 
+\bibitem{heaton01} Heaton, J., \textbf{\textit{Encog Machine Learning Framework}}, http://heatonresearch.com/encog/
+
 \bibitem{hristov01} Hristov, P., Dimitrov, W., \textbf{\textit{The blockchain as a backbone of GDPR compliant frameworks}}, Proceedings of 8th International Multidisciplinary Symposium - Challenges and Opportunities For Sustainable Development Through Quality and Innovation in Engineering and Research Management, \textbf{20}(1), p.
@@
 458--467, 2005.
+
+\bibitem{balabanov01} Balabanov, T., \textbf{\textit{VitoshaTrade Project - vitoshatrade.veldsoft.eu}}, http://github.com/VelbazhdSoftwareLLC/VitoshaTrade
 

### 2019-09-13 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/paper.tex
Some fixes were done.

--- 2019-09-05 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/paper.tex
@@
 One of the most famous donated distributed computing project is SETI@home, which is related to deep space signals processing in attempt to find an alien life.
-The usage of donated distributed computing had its influence in the time series forecasting in the face of MoneyBee\cite{bohn01} project.
+The usage of donated distributed computing had its influence in the time series forecasting in the face of MoneyBee \cite{bohn01} project.
 MoneyBee project was a desktop screensaver application, which was calculating financial time series forecasting by training of artificial neural networks with evolutionary algorithms.
 With the expansion of the mobile devices in the last decade it become relevant some donated distributed computing solutions to be developed as mobile applications.
-Such a solution was developed at IICT-BAS\cite{tomov01}, which is based on Android Live Wallpaper technology.
+Such a solution was developed at IICT-BAS \cite{tomov01}, which is based on Android Live Wallpaper technology.
 This research proposes an extension of the work done at IICT-BAS in the direction of human-computer based distributed computing by providing software capabilities of the users to vote for future financial changes.
@@
 In some studies evolutionary algorithms are used for artificial neural network topology optimization \cite{kapanova01}, but in other studies evolutionary algorithms are used for artificial neural network weights optimization \cite{aljarah01}.
-When evolutionary algorithms are used for weights optimization it is very common the training to be combined with back-propagation algorithm.
+When evolutionary algorithms are used for weights optimization it is very common the training to be combined with back-propagation algorithm \cite{balabanov02}.
 Even when such a hybrid training algorithm is used it is in the class of supervised training.
@@
 Evolutionary algorithms have a common advantage compared to the exact numerical methods and that is the possibility for parallel calculations \cite{altinoz01}.
-When the parallel calculations are done on separate mobile devices it is the case of mobile distributed computing.
+When the parallel calculations are done on separate mobile devices it is the case of mobile distributed computing \cite{sharma01}.
 It is very common the financial forecasting to be done by the usage of the past values in the time series, but it can be also combined with human evaluation of the future values.
@@
 The most successful forecast votes of the users are used in artificial neural network training process.
+
+This paper is organized as follows.
+Section 1 gives an overview of the problem  with a special emphasis on ANNs/GAs and their strengths/weaknesses when applied for the problem's solution.
+Section 2 introduces a distributed computing system based on mobile devices with the extension to human-computer distributed computing.
+Experiments and results are presented in Section 3.
+The final Section 4 concludes and some further work suggestions are provided.
 
@@
 The main advantage of this approach is that it is using donated computation power of many participants.
-In the same time the biggest disadvantage is that the screensaver is used relatively rare, only when the user is not using the computer.
+In the same time the biggest disadvantage is that the screensaver is used relatively rare, only when the user is not using his/her computer.
 The other disadvantage is that personal computers are not used 24/7.
@@
 \ref{fig02}), developed in IICT-BAS.
-The implementation is done under an Android operating system as active wallpaper.
+The implementation is done under Android operating system as active wallpaper.
 Inside the forecasting module a multilayer perceptron is used.
 The training algorithm is a combination between back-propagation and differential evolution.
-All calculations are done as active Android application, which works 24/7 in a background mode.
+All calculations are done as active wallpaper Android application, which works 24/7 in a background mode.
 When better artificial neural network weights are found a communication with a remote server is established.
@@
 
-In both projects only a pure mathematical information is used for prediction calculations.
-This limitation was overcome in ElectricSheep\cite{draves01} project.
-In this project it is used a rating system whereat the users vote for different fractal generated animations.
+In both projects only a pure mathematical information is used for forecasts calculations.
+Such limitation was overcome in ElectricSheep \cite{draves01} project.
+In ElectricSheep a rating system whereat the users vote for different fractal generated animations is used.
 Voting is done in two directions - thumb up and thumb down.
@@
 The rating is not constant and it goes down with the time passing.
+In the proposed model direction voting is implemented, just like it was done in ElectricSheep project.
+Users vote only up if they think that the price will go up or down if they think that the price will go down.
+There is no limit how many times the user will press up or down button, just like it is implemented in ElectricSheep project.
+The intensity of the single user voting in one time interval (daily intervals for example) does not count.
+Only the first push of up/down button is taken in account.
+Users' voting is irregular, because in real-time operation the operator of the system has no control over donated human-computer participation.
 
 This study proposes a combination of the projects described above.
-The VitoshaTrade project is extended in the direction of voting system addition.
+The VitoshaTrade \cite{balabanov01} project is extended in the direction of voting system addition.
 The Android operating system offers software modules called widgets.
@@
 
-Voting of the users is collected on PHP/MySQL based web server\cite{tomov01}, a part of the server side implementation of the VitoshaTrade project.
+Voting of the users is collected on PHP/MySQL based web server \cite{tomov01}, as part of the server side implementation of the VitoshaTrade project.
 On the server side additional analysis is done.
@@
 A back-propagation combined with differential evolution training of multilayer perceptron is done.
+As topology for the multilayer perceptron 7-5-3 is used implemented with Encog Machine Learning Framework \cite{heaton01} for Java.
+The input has 7 valuse, the output has 3 values and the hidden layer was seledted to be $5 = (7 + 3) / 2$.
 The training process is continuous, which means that the forecasting is done parallel with the training.
@@
 \begin{figure}
-\includegraphics[width=0.8\textwidth]{fig03.png}
+\includegraphics[width=1.0\textwidth]{fig03.png}
 \centering
-\caption{Forex EUR/USD time series.} \label{fig03}
+\caption{ForEx EUR/USD time series.} \label{fig03}
 \end{figure}
 \FloatBarrier
+
+There is no limit how many people can participate in the proposed human-computer distributed computing system.
+The system itself is highly scalable, but experiments in this study were done in a closed group of the team which is working on the research.
 
@@
 
-As a future research it will be interesting different communication capabilities to be investigated as described in \cite{alexandrov01}.
+As a future research it will be interesting different communication capabilities to be investigated as described in \cite{alexandrov01} and working proceses optimization as described in \cite{atanasov01}.
 Also it will be interesting generalized artificial neural networks \cite{tashev01} to be used instead of multilayer perceptron.
@@
 \begin{thebibliography}{8}
+\bibitem{bohn01} Bohn, A., Guting, T., Mansmann, T., \textbf{\textit{MoneyBee: A new product to predict stock market developments using artificial intelligence and increased calculation capacitiy}} (in German), et al.
+Wirtschaftsinf, \textbf{45}(3), p.
+325--333, 2003.
+
+\bibitem{tomov01} Tomov, P., Zankinski, I., Barova, M., \textbf{\textit{Mobile Alternative of the MoneyBee Project for Financial Forecasting}}, Proceedings of the Annual University Scientific Conference of the National Military University Vasil Levski, Veliko Tarnovo, p.
+1085--1089, 2018.
+
 \bibitem{nava01} Nava, N., Di Matteo, T., Aste, T., \textbf{\textit{Financial Time Series Forecasting Using Empirical Mode Decomposition and Support Vector Regression}}, RISKS, \textbf{6}(1), article 7, 2018.
@@
 
+\bibitem{balabanov02} Balabanov, T., Genova, K., \textbf{\textit{AJAX Distributed System for Evolutionary Algorithms based Artificial Neural Networks Training}}, XXIV International Symposium on Management of Thermal Energy Facilities and Systems, Management of Energy, Industrial and Environmental Systems, John Atanasoff Union of Automation and Informatics, p.
+49--52, 2016.
+
 \bibitem{altinoz01} Altinoz, O.T., Deb, K., \textbf{\textit{Late parallelization and feedback approaches for distributed computation of evolutionary multi-objective optimization algorithms}}, Neural Computing and Applications, Springer London, \textbf{30}(3), p.
@@
 
-\bibitem{bohn01} Bohn, A., Guting, T., Mansmann, T., \textbf{\textit{MoneyBee: A new product to predict stock market developments using artificial intelligence and increased calculation capacitiy}} (in German), et al.
-Wirtschaftsinf, \textbf{45}(3), p.
-325--333, 2003.
+\bibitem{sharma01} Sharma, R., Rodriguez, T., \textbf{\textit{Distributed computing for portable computing devices}}, US10282470B2, United States, 2019.
 
-\bibitem{tomov01} Tomov, P., Zankinski, I., Barova, M., \textbf{\textit{Mobile Alternative of the MoneyBee Project for Financial Forecasting}}, Proceedings of the Annual University Scientific Conference of the National Military University Vasil Levski, Veliko Tarnovo, p.
-1085--1089, 2018.
+\bibitem{heaton01} Heaton, J., \textbf{\textit{Encog Machine Learning Framework}}, http://heatonresearch.com/encog/
 
@@
 
+\bibitem{balabanov01} Balabanov, T., \textbf{\textit{VitoshaTrade Project - vitoshatrade.veldsoft.eu}}, http://github.com/VelbazhdSoftwareLLC/VitoshaTrade
+
 \bibitem{alexandrov01} Alexandrov, A., \textbf{\textit{Comparative analysis of IEEE 802.15.4 based communication protocols used in wireless intelligent sensor systems}}, Proceedings of the International conference RAM, p.
 51--54, 2014.
+
+\bibitem{atanasov01} Atanasov, J., Atanasova, T., \textbf{\textit{Optimizing the management and control of apparel enterprise by information technologies}}, 19th Telecommunications Forum, TELFOR 2011 - Proceedings of Papers, Article number 6143777, p.
+1245--1248, Belgrade, Serbia, 2011.
 

### 2019-09-14 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/paper.tex
Some more fixes were done.

@@
 Financial time series forecasting is an attractive and intensively researched area \cite{nava01}.
+Financial forecasting has two common components - direction of the change and magnitude of the change.
+In many situations it is enough direction of the change to be predicted.
+After that decision makers can do their decisions.
+If magnitude of the change is also predicted decisions taken would be much more accurate.
+To predict the magnitude of the change is relevant to predict the direction of the next serious change in the opposite direction.
 One of the most interesting research directions in this field is related to the usage of artificial neural networks combined with evolutionary algorithms \cite{zhang01}.
@@
 
-\begin{figure}
-\includegraphics[width=1.0\textwidth]{fig02.png}
-\centering
-\caption{VitoshaTrade active wallpaper} \label{fig02}
-\end{figure}
-\FloatBarrier
-
 In both projects only a pure mathematical information is used for forecasts calculations.
@@
 In the proposed model direction voting is implemented, just like it was done in ElectricSheep project.
-Users vote only up if they think that the price will go up or down if they think that the price will go down.
+Users vote only up if they think that the price will go up or down if they think that the price will go down (Fig.
+\ref{fig04}).
 There is no limit how many times the user will press up or down button, just like it is implemented in ElectricSheep project.
@@
 Users' voting is irregular, because in real-time operation the operator of the system has no control over donated human-computer participation.
+
+\begin{figure}
+\includegraphics[width=1.0\textwidth]{fig02.png}
+\centering
+\caption{VitoshaTrade active wallpaper} \label{fig02}
+\end{figure}
+\FloatBarrier
 
@@
 In this study a composite widget is proposed, which shows the thicker of the currency pair (EUR/USD in the example provided), the value of the rate between the two currencies, and arrows for vote up or vote down.
+
+\begin{figure}
+\includegraphics[width=1.0\textwidth]{fig04a.png}
+\centering
+\caption{Voting widget positioned over active wallpaper} \label{fig04}
+\end{figure}
+\FloatBarrier
 

### 2019-09-15 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/paper.tex
Final fixes were done.

--- 2019-09-13 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/paper.tex
@@
 \begin{abstract}
-Distributed computing became very popular in the last two decades.
-In many cases distributed computing projects are based on a donated calculation power.
-One of the most famous donated distributed computing project is SETI@home, which is related to deep space signals processing in attempt to find an alien life.
-The usage of donated distributed computing had its influence in the time series forecasting in the face of MoneyBee \cite{bohn01} project.
-MoneyBee project was a desktop screensaver application, which was calculating financial time series forecasting by training of artificial neural networks with evolutionary algorithms.
+Distributed computing became very popular in the last two decades and in many cases such projects are based on a donated calculation power.
 With the expansion of the mobile devices in the last decade it become relevant some donated distributed computing solutions to be developed as mobile applications.
-Such a solution was developed at IICT-BAS \cite{tomov01}, which is based on Android Live Wallpaper technology.
+Such a solution was developed at IICT-BAS \cite{tomov01}, which is based on an Android Live Wallpaper technology.
 This research proposes an extension of the work done at IICT-BAS in the direction of human-computer based distributed computing by providing software capabilities of the users to vote for future financial changes.
+At the beginning, a brief overview of the topic with special emphasis on ANNs/GAs and their strengths/weaknesses when applied for the problem's solution is given.
+After that the study treats the aspect of extension of a distributed computing system based on mobile devices to human-computer distributed computing.
+Experiments and results are presented in Section 3 and the final Section 4 concludes and provides some further work suggestions.
 
@@
 
-Financial time series forecasting is an attractive and intensively researched area \cite{nava01}.
+One of the most famous donated distributed computing project is SETI@home, which is related to deep space signals processing in attempt to find an alien life \cite{setihome01}.
+But donated distributed computing has its influence also in other areas like the time series forecasting.
+The MoneyBee project was a desktop screensaver application, which was calculating financial time series forecasting by training of artificial neural networks with evolutionary algorithms \cite{bohn01}.
+Financial time series forecasting is an attractive and intensively researched area \cite{nava01}, whereat financial forecasting has two common components - direction of the change and magnitude of the change.
+In many situations it is enough direction of the change to be predicted.
+After that decision makers are able to do their decisions.
+If magnitude of the change is also predicted decisions taken would be much more accurate and to be able to predict the magnitude of the change is crucial to predict the direction of the next serious change in the opposite direction.
+
 One of the most interesting research directions in this field is related to the usage of artificial neural networks combined with evolutionary algorithms \cite{zhang01}.
@@
 When evolutionary algorithms are used for weights optimization it is very common the training to be combined with back-propagation algorithm \cite{balabanov02}.
-Even when such a hybrid training algorithm is used it is in the class of supervised training.
+Even when such a hybrid training algorithm is used, it is in the class of supervised training.
 Training examples are fed to the artificial neural network and its weights are modified according the selected training algorithm.
-Evolutionary algorithms have a common advantage compared to the exact numerical methods and that is the possibility for parallel calculations \cite{altinoz01}.
+Generally, evolutionary algorithms have a common advantage compared to the exact numerical methods and that is the possibility for parallel calculations \cite{altinoz01}.
 When the parallel calculations are done on separate mobile devices it is the case of mobile distributed computing \cite{sharma01}.
 It is very common the financial forecasting to be done by the usage of the past values in the time series, but it can be also combined with human evaluation of the future values.
+
 In this research a human-computer based mobile distributed computing solution is proposed.
@@
 
-This paper is organized as follows.
-Section 1 gives an overview of the problem  with a special emphasis on ANNs/GAs and their strengths/weaknesses when applied for the problem's solution.
-Section 2 introduces a distributed computing system based on mobile devices with the extension to human-computer distributed computing.
-Experiments and results are presented in Section 3.
-The final Section 4 concludes and some further work suggestions are provided.
-
 \section{Model Proposition} \label{Model Proposition}
 
-MoneyBee screensaver (Fig.
-\ref{fig01}) was organized as a desktop application, which runs in background when the computer is in idle mode.
-Artificial neural networks were trained with the usage of evolutionary algorithms.
-The main advantage of this approach is that it is using donated computation power of many participants.
+As already mentioned above, a famous example in the area of artificial neural network is the project MoneyBee \cite{bohn01}.
+The MoneyBee screensaver (Fig.
+\ref{fig01}) was organized as a desktop application, which runs in background when the computer is in idle mode and the artificial neural networks were trained with the usage of evolutionary algorithms.
+The main advantage of this approach is the usage of donated computation power of many participants.
 In the same time the biggest disadvantage is that the screensaver is used relatively rare, only when the user is not using his/her computer.
-The other disadvantage is that personal computers are not used 24/7.
+The other disadvantage is that personal computers are not in operation 24/7.
 
@@
 These disadvantages are addressed in the VitoshaTrade project (Fig.
-\ref{fig02}), developed in IICT-BAS.
-The implementation is done under Android operating system as active wallpaper.
+\ref{fig02}), which is developed in IICT-BAS (Institute of Information and Communication Technologies - Bulgarian Academy of Sciences) \cite{balabanov01}.
+The implementation is done under Android operating system as an active wallpaper.
 Inside the forecasting module a multilayer perceptron is used.
 The training algorithm is a combination between back-propagation and differential evolution.
-All calculations are done as active wallpaper Android application, which works 24/7 in a background mode.
+An active wallpaper Android application executes all calculations and it works 24/7 in a background mode.
 When better artificial neural network weights are found a communication with a remote server is established.
 Active wallpapers are used to visualize information on a full screen behind all other visual components.
+
+In both projects only a pure mathematical information is used for forecasts calculations.
+Such limitation was overcome in ElectricSheep \cite{draves01} project.
+In ElectricSheep is implemented a rating system whereat the users vote for different fractal generated animations.
+Voting is done in two directions - thumb up and thumb down.
+Each vote increases or decreases individual rating, which is directly related to its survival chances.
+The rating is not constant and it goes down with the time passing.
+In the proposed model a direction voting is implemented similar to the ElectricSheep project.
+Users vote only up if they think that the price will go up, and down if they think that the price will drop (Fig.
+\ref{fig04}), whereat there is no limit how many times the user will press up or down buttont.
+The intensity of the single user voting in one time interval (daily intervals for example) does not count.
+Only the first push of up/down button is taken in account.
+Users' voting is irregular, because in real-time operation the operator of the system has no control over donated human-computer participation.
 
@@
 
-In both projects only a pure mathematical information is used for forecasts calculations.
-Such limitation was overcome in ElectricSheep \cite{draves01} project.
-In ElectricSheep a rating system whereat the users vote for different fractal generated animations is used.
-Voting is done in two directions - thumb up and thumb down.
-Each vote increases or decreases individual rating, which is directly related to its survival chances.
-The rating is not constant and it goes down with the time passing.
-In the proposed model direction voting is implemented, just like it was done in ElectricSheep project.
-Users vote only up if they think that the price will go up or down if they think that the price will go down.
-There is no limit how many times the user will press up or down button, just like it is implemented in ElectricSheep project.
-The intensity of the single user voting in one time interval (daily intervals for example) does not count.
-Only the first push of up/down button is taken in account.
-Users' voting is irregular, because in real-time operation the operator of the system has no control over donated human-computer participation.
-
 This study proposes a combination of the projects described above.
-The VitoshaTrade \cite{balabanov01} project is extended in the direction of voting system addition.
-The Android operating system offers software modules called widgets.
-Widgets have graphical user interface and they occupy small part of the visible working area.
-Most commonly widgets are used for representation of small pieces of information, such astronomical time, weather forecast, calendar and many others.
+The VitoshaTrade \cite{balabanov01} project is extended in the direction of additional voting system.
+The Android operating system offers software modules called widgets, which have graphical user interface and they occupy small part of the visible working area.
+Most commonly, widgets are used for representation of small pieces of information, such astronomical time, weather forecast, calendar, and others.
 Widgets are much more interactive than the active wallpaper.
-The user is capable to input small pieces of information as voting for example.
-The other better advantage of the widgets is that users have much better control over visualization and placement of the widgets.
+The user is capable to input small pieces of information as voting, for example.
+Another advantage of the widgets is that users have much better control over visualization and placement of the widgets.
 In this study a composite widget is proposed, which shows the thicker of the currency pair (EUR/USD in the example provided), the value of the rate between the two currencies, and arrows for vote up or vote down.
 
-Voting of the users is collected on PHP/MySQL based web server \cite{tomov01}, as part of the server side implementation of the VitoshaTrade project.
-On the server side additional analysis is done.
+\begin{figure}
+\includegraphics[width=1.0\textwidth]{fig04a.png}
+\centering
+\caption{Voting widget positioned over active wallpaper} \label{fig04}
+\end{figure}
+\FloatBarrier
+
+Voting of the users is collected on PHP/MySQL based web server \cite{tomov01} as part of the server side implementation of the VitoshaTrade project.
+On the server side additional analysis are done.
 Votes which were more accurate are better taken for next generations in the global distributed differential evolution population.
@@
 \ref{fig03}).
-Time series is disassembled in training examples (7 values for lag and 3 values for lead).
-A back-propagation combined with differential evolution training of multilayer perceptron is done.
-As topology for the multilayer perceptron 7-5-3 is used implemented with Encog Machine Learning Framework \cite{heaton01} for Java.
-The input has 7 valuse, the output has 3 values and the hidden layer was seledted to be $5 = (7 + 3) / 2$.
+Time series is disassembled in training examples (7 values for lag and 3 values for lead) and a back-propagation combined with differential evolution training of multilayer perceptron is additionally done.
+As topology for the multilayer perceptron 7-5-3 is used an implementation with Encog Machine Learning Framework \cite{heaton01} for Java.
+The input has value 7, the output has value 3 and the hidden layer is selected to be $5 = (7 + 3) / 2$.
 The training process is continuous, which means that the forecasting is done parallel with the training.
@@
 More reliable votes are taken with higher coefficient for the further predictions.
-According to some regulations, like GDPR in EU \cite{hristov01}, a special care should be taken for the personal information collected on the servers side.
+According to some regulations, like GDPR (General Data Protection Regulation) in EU (European Union) \cite{hristov01}, a special care should be taken for the personal information collected on the servers side.
 
@@
 There is no limit how many people can participate in the proposed human-computer distributed computing system.
-The system itself is highly scalable, but experiments in this study were done in a closed group of the team which is working on the research.
+The system itself is highly scalable, but experiments in this study were done in a closed group of the team by IICT-BAS (8 users), which is working on the research.
 
@@
 Values measured in time are just one projection of the process which goes under it.
-Any computer based calculation nowadays is not capable to act as human intuition.
+Any computer based calculation nowadays is not capable to act like human intuition.
 Involving human intuition in the forecasting process is the main advantage of the proposed model.
-The mechanisms behind human intuition are not investigated enough and are still very difficult to recreate in the modern computers.
+The mechanisms behind human intuition are not investigated enough and they are still very difficult to be recreated in the modern computers.
 Because of these reasons the proposed financial time series forecasting solution advances the pure technical analysis with minimal influence of human impact.
-The proposed simple voting system consolidated the power of machine learning algorithms and the human sixth sense feeling in an efficient and cost effective financial forecasting infrastructure.
+The proposed simple voting system consolidates the power of machine learning algorithms and the human sixth sense feeling in an efficient and cost effective financial forecasting infrastructure.
 
-As a future research it will be interesting different communication capabilities to be investigated as described in \cite{alexandrov01} and working proceses optimization as described in \cite{atanasov01}.
+As a future research it will be interesting to investigate different communication capabilities as described in \cite{alexandrov01} and working processes optimization as described in \cite{atanasov01}.
 Also it will be interesting generalized artificial neural networks \cite{tashev01} to be used instead of multilayer perceptron.
+More attention shoud be paid to the security, how to increase the dependability of distributed systems by providing better protection against deliberate attacks.
+Higher reliability, ubiquity, and better security are only some aspects, which in further research are desired.
 
@@
 
-\bibitem{balabanov01} Balabanov, T., \textbf{\textit{VitoshaTrade Project - vitoshatrade.veldsoft.eu}}, http://github.com/VelbazhdSoftwareLLC/VitoshaTrade
+\bibitem{balabanov01} Balabanov, T., \textbf{\textit{VitoshaTrade Project - vitoshatrade.veldsoft.eu}}, http://github.com/VelbazhdSoftwareLLC/VitoshaTrade.
 
@@
 
-\bibitem{tashev01} Tashev, T., Hristov, H., \textbf{\textit{Modeling of synthesis of information processes with generalized nets}}, Cybernetics and Information Technologies, \textbf{3}(2), p.
+\bibitem{tashev01} Tashev, T., Hristov, H., \textbf{\textit{MoneyBee: Aktienkursprognose mit kuenstlicher intelligenz bei hoher rechenleistung}}, Wirtschaftsinformatik, \textbf{3}(2), p.
 92--104, 2003.
+
+\bibitem{setihome01}University of California, \textbf{\textit{SETI@home}}, https://setiathome.berkeley.edu/.
+
 \end{thebibliography}

### 2019-09-24 Human-Computer-Mobile-Distributed-Computing-for-Time-Series-Forecasting/paper.tex
Small errors were fixed.

@@
 	Plamen Petrov \and \\
-	Iliyan Zankinskir \and
+	Iliyan Zankinski \and
 	Todor Balabanov\orcidID{0000-0003-3139-069X}
@@
 \authorrunning{R. Ketipov et al.}
+\titlerunning{Human-Computer MDC for TS Forecasting}
 
